---
title: "Project Question 1"
format: html
---

```{r}
# Load necessary libraries
options(warn=-1)
if (!require("pacman")) 
  install.packages("pacman")

# use this line for installing/loading
pacman::p_load(tidyverse,
               glue,
               scales,
               openintro,
               gridExtra,
               ggrepel,
               ggmap,
               ggridges,
               dsbox,
               devtools,
               fs,
               janitor,
               here,
               dplyr) 
  
```

## Plan out layer mappings

The following libraries / functions will be used in order to answer Question 1, "How have the frequency and depth of seismic events at Mount Vesuvius changed over time?"

-   geom_point:

```         
-   Creating a scatter plot of the seismic events (depth and duration-magnitude) over time
```

-   geom_line:

    -   Creating a scatter plot of the seismic events (depth and duration-magnitude) over time

-   geom_bar:

    -   View the frequency of the seismic events aggregated per year or year, month

```{r}
vesuvius_data <- read.csv(here("data","vesuvius.csv"))
vesuvius_data <- vesuvius_data |>
    mutate(
        date = str_replace_all(time, "T", " "),
        date = str_replace_all(date, "Z", ""),
        date = as.POSIXct(date, format = "%Y-%m-%d %H:%M:%S")
    )
vesuvius_data_filtered <- vesuvius_data %>% filter(!is.na(duration_magnitude_md))
glimpse(vesuvius_data)
glimpse(vesuvius_data_filtered)

```

```{r}
vesuvius_data_filtered$year <- format(vesuvius_data_filtered$date, "%Y")
vesuvius_data_filtered$month <- format(vesuvius_data_filtered$date, "%m")
vesuvius_data_filtered$year <- as.numeric(as.character(vesuvius_data_filtered$year))
vesuvius_data_filtered$year <- as.numeric(as.character(vesuvius_data_filtered$year))
```

```{r}
aggregated_data <- vesuvius_data_filtered %>%
  group_by(year, month) %>%
  summarise(mean = mean(duration_magnitude_md))
```

```{r}
aggregated_data$date <- paste0(aggregated_data$year, " ", aggregated_data$month)
glimpse(aggregated_data)
```

```{r}
ggplot(aggregated_data, 
            aes(x = year, y = mean, color = year)) +
            geom_jitter() +
  geom_smooth(method = "loess", 
              se = FALSE)
```

## Plan out data wrangling methods

This section summarizes the data wrangling completed:

-   time:

    -   Replace "T" with " "

    -   Remove "Z"

    -   Convert into a date variable

    -   Extract the years

    -   Extract the months